Highly Available Transactions: Virtues and Limitations

### 2013-04-01 writeup/hat-analysis.tex
spell

\vspace{.5em}\noindent\textit{{Writes Follow Reads}} requires that, if
a session observes an effect of transaction $T_1$ and subsequently
commits transaction $T_2$, then another session can only observe
effects of $T_2$ if it can also observe $T_1$'s effects (or later
values that supercede $T_1$'s).  Any order on transactions should
respect the reads-from order.\vspace{.5em}

@@

It is impossible to prevent Lost Update in a highly available
environment. Consider two clients who submit the following $T_1$ and
$T_2$ on opposite sides of a network partition:
\begin{align*}
\small
T_1 &: r_x(100)~w_x(100+20=120)
\\T_2 &: r_x(100)~w_x(100+30=130)
\end{align*}
Regardless of whether $x=120$ or $x=130$ is chosen by a replica, the
database state could not have arisen serial execution of $T_1$ and
$T_2$.\footnote{In this example, we assume that, as is standard in
  modern databases, replicas accept values as they are written (i.e.,
  register semantics). This particular example could be made
  serializable via the use of commutative updates
  (Section~\ref{sec:evaluation}) but, as we show in the extended version of
  the paper~\cite{hat-tr}, the problem persists in the general case.}
To prevent this from happening, either $T_1$ or $T_2$ should not have
committed. Each client's respective server might try to detect that
another write occurred, but this requires knowing the version of the
latest write to $x$. In our example, this reduces to a requirement for
linearizability, which is, via Gilbert and Lynch's proof of the CAP
Theorem, provably unachievable with high
availabilty~\cite{gilbert-cap}.

>>>>

\vspace{.5em}\noindent\textit{{Writes Follow Reads}} requires that, if
a session observes an effect of transaction $T_1$ and subsequently
commits transaction $T_2$, then another session can only observe
effects of $T_2$ if it can also observe $T_1$'s effects (or later
values that supersede $T_1$'s).  Any order on transactions should
respect the reads-from order.\vspace{.5em}

@@

It is impossible to prevent Lost Update in a highly available
environment. Consider two clients who submit the following $T_1$ and
$T_2$ on opposite sides of a network partition:
\begin{align*}
\small
T_1 &: r_x(100)~w_x(100+20=120)
\\T_2 &: r_x(100)~w_x(100+30=130)
\end{align*}
Regardless of whether $x=120$ or $x=130$ is chosen by a replica, the
database state could not have arisen serial execution of $T_1$ and
$T_2$.\footnote{In this example, we assume that, as is standard in
  modern databases, replicas accept values as they are written (i.e.,
  register semantics). This particular example could be made
  serializable via the use of commutative updates
  (Section~\ref{sec:evaluation}) but, as we show in the extended version of
  the paper~\cite{hat-tr}, the problem persists in the general case.}
To prevent this from happening, either $T_1$ or $T_2$ should not have
committed. Each client's respective server might try to detect that
another write occurred, but this requires knowing the version of the
latest write to $x$. In our example, this reduces to a requirement for
linearizability, which is, via Gilbert and Lynch's proof of the CAP
Theorem, provably unachievable with high
availability~\cite{gilbert-cap}.
